Ship-based lidar evaluation of Southern Ocean low clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses

### 2025-06-30 manuscript_agu_rev1.tex
Addressing author comments

\begin{keypoints}
\item Ground-based lidar evaluation of a km-scale global climate model and reanalyses reveals substantial cloud biases over the Southern Ocean.
\item Total cloud fraction and fog are underestimated in the reanalyses. In all models, a cloud peak at 500~m is overestimated and too high.
\item A ``too few, too bright'' problem of underestimated cloud fraction, compensated by overestimated cloud albedo, is present in MERRA-2.
\end{keypoints}

@@

@@

We partitioned our data into two mutually exclusive subsets by cyclonic activity. For this purpose, we used a cyclone tracking algorithm to identify extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis and ICON data. We used the open-source cyclone tracking package CyTRACK \cite{perez-alarcon2024}. Generally, what constitutes an EC is considered relatively arbitrary due to the very large variability of ECs \cite{neu2013}. The CyTRACK algorithm uses mean sea level pressure and wind speed thresholds as well as tracking across time steps to identify cyclone centers and radii in each time step. With this information, we could classify every location at a given time as either cyclonic or non-cyclonic. Due to a relatively small total area covered by cyclones (as identified by the cyclone center and radius), we chose a circle of double the radius (relative to one identified by CyTRACK) centered at the cyclone center as a cyclonic area for every time step and cyclone. All other areas were identified as non-cyclonic. For identifying cyclones in the observations and the reanalyses, ERA5 pressure and wind fields were used as the input to CyTRACK. This is justified by the fact that the large-scale pressure and wind fields in ERA5 are likely sufficiently close to reality. \citeA{mcerlich2023} have shown that wind is simulated well in ERA5 relative to the WindSat polarimetric microwave radiometer measurements \cite{meissner2009}. For identifying cyclones in ICON, its own pressure and wind fields were used as the input to CyTRACK because the model is free-running, and thus the pressure and wind fields are different from reality. Subsetting by proximity to cyclones is a relatively crude measure because it does not take into account the different sectors of cyclones, which are commonly associated with different weather situations. However, this was a choice made for simplicity of the analysis, given the quantity of data. \citeA{konstali2024} performed a more complex attribution of precipitation to individual cyclone features.

>>>>

\begin{keypoints}
\item Large cloud biases are present in the global storm-resolving model, but are lower in several key aspects than in the reanalyses.
\item The models tend to underestimate total cloud fraction, near-surface cloud and fog, while overestimating cloud occurrence peak at 500 m.
\item A ``too few, too bright'' problem of underestimated cloud fraction, compensated by overestimated cloud albedo, is present in MERRA-2.
\end{keypoints}

@@

Some of the observatial data were likely used in reanalysis assimilation. The Macquarie Island station surface measurements and radiosonde profiles (not used in our analysis), were sent to the World Meteorological Organization Global Telecommunication System (GTS). The lidar measurements and radiosonde profiles on the RSV \textit{Aurora Australis} were not sent for assimilation.

@@

We partitioned our data into two mutually exclusive subsets by cyclonic activity. For this purpose, we used a cyclone tracking algorithm to identify extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis and ICON data. We used the open-source cyclone tracking package CyTRACK \cite{perez-alarcon2024}. Generally, what constitutes an EC is considered relatively arbitrary due to the very large variability of ECs \cite{neu2013}. The CyTRACK algorithm uses mean sea level pressure and wind speed thresholds as well as tracking across time steps to identify cyclone centers and their radii in each time step. With this information, we could classify every location at a given time as either cyclonic or non-cyclonic. Due to a relatively small total area covered by cyclones, as identified by the cyclone center and radius, for every time step and cyclone we defined a cyclonic area as a circle of double the radius identified by CyTRACK centered at the cyclone center. All other areas were defined as non-cyclonic. For identifying cyclones in the observations and the reanalyses, ERA5 pressure and wind fields were used as the input to CyTRACK. This is justified by the fact that the large-scale pressure and wind fields in ERA5 are likely sufficiently close to reality. \citeA{mcerlich2023} have shown that wind is simulated well in ERA5 relative to the WindSat polarimetric microwave radiometer measurements \cite{meissner2009}. For identifying cyclones in ICON, its own pressure and wind fields were used as the input to CyTRACK because the model is free-running, and thus the pressure and wind fields are different from reality. Subsetting by proximity to cyclones is a relatively crude measure because it does not take into account the different sectors of cyclones, which are commonly associated with different weather situations. However, this was a choice made for simplicity of the analysis, given the quantity of data. \citeA{konstali2024} performed a more complex attribution of precipitation to individual cyclone features.
